---
title: "Hierarchical Clustering"
output: html_notebook
---

Version 1.0, July 2025, SA

This script clusters the clusters resulting from a long-axis x age bin interaction into superclusters that share similar profiles of connectivity with the anteroposterior hippocampus (as visualized in Figure 4A)

It also plots several cluster quality checks (including supplemental plots in Fig.S1.B-E)

Input: Hippo_BinxAxis_F_betas.txt

# Packages and functions
```{r}

library(tidyr)
library(factoextra)
library(fpc)
library(cluster)
library(cowplot)
library(nlme)
library(ggplot2)
library(Rmisc)
library(emmeans)

```
# Cluster the Clusters

## Read in and organize the data 
```{r}

## read in the data. CHANGE TO YOUR DIRECTORY. 
dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos"
data <- read.table(paste(dir, "/Hippo_BinxAxis_F_betas.txt", sep=""), header = TRUE) 

### use top 44 for cluster threshold of 10
data<-subset(data, clust < 45)

## average values to cluster, across subj and hem. 
data_ag<-aggregate(data[, "beta"], by=(list(data$ax, data$bin, data$clust)), mean)
colnames(data_ag)<-c("ax","bin","clust","beta")

## add a column that is ax x bin 
data_ag$cond<-paste(data_ag$ax, data_ag$bin, sep="_bin")

## drop the other columns 
data_ag2<-data.frame(data_ag[,c("clust","beta","cond")])

## wide format, want one row per cluster. Betas in the cells. 
data_clust<-spread(data_ag2, cond, beta)

## keep clust as its on label so know the order, but drop it so doesnt get included in the cluster analysis 
data_cluster_labels<-data.frame(data_clust$clust)

### subtract ant post to get a difference score to cluster
bin1<-data_clust$ant_bin1 - data_clust$post_bin1
bin2<-data_clust$ant_bin2 - data_clust$post_bin2
bin3<-data_clust$ant_bin3 - data_clust$post_bin3
bin4<-data_clust$ant_bin4 - data_clust$post_bin4

data_ap<-cbind(bin1,bin2,bin3,bin4)

```

## Hierarchical clustering 
```{r}
## scale it, as variables with different scales have differing influence over the calculation of distance
data_clust2<-data.frame(scale(data_ap))

## add cluster labels to data_ap
data_ap<-cbind(data_cluster_labels, data_ap)

## euclidean distance matrix on the version without the cluster labels 
distance.matrix <- dist(data_clust2, method="euclidean")

## cluster based on ward's method 
hierarchical.cluster <- hclust(distance.matrix)

```

## Save cluster solution 

We chose k=6 based on the agglomeration schedule + diagnostic plots below 

```{r}

data_cluster_labels<- data.frame(data_cluster_labels, clusters=cutree(hierarchical.cluster, k=6))
#write.csv(data_cluster_labels, file = "data_cluster_labels.csv", row.names = FALSE)

data_ap <- data.frame( data_ap, clusters=cutree(hierarchical.cluster, k=6))

```

# Clustser Quality Checks 

## determine how clusterable 
```{r}
## determine clusterability 
res <- get_clust_tendency(data_clust2, n = nrow(data_clust2)-1, graph = FALSE)
res$hopkins_stat

```

## Plot agglomeration schedule 
```{r}

#png(file="~/Downloads/test.png", width=2000, height=2000, res=300)
#Print a reversed Scree plot
plot(hierarchical.cluster$height, ylab="height", xlab="agglomeration step", main="agglomeration schedule", 
     cex=2, cex.lab=2, cex.axis=2, cex.main=2.5) ## looks like k=6
#dev.off()

#Print a table similar to an Agglomeration Schedule
cbind(hierarchical.cluster$merge, hierarchical.cluster$height)

```

## Plot dendrogram
```{r}

 fviz_dend(hierarchical.cluster, k=6, k_colors = c("#18e0f2","#152370","#f5e107","#f5200c","#fa750f","#4bf52a"), as.ggplot=TRUE) +
 theme(plot.title = element_text(size=24, face="bold", hjust=0.5), axis.ticks.x=element_blank()) +
  theme(axis.title = element_text(size=20, hjust=0.5), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=16)) +
  #scale_y_continuous(limits = c(-2,10), breaks=seq(0,10,2))+
  ggtitle("cluster dendrogram") +
  labs(x="", y="height")

#ggsave("dendrogram.png", path="~/Downloads", width = 6.5, height = 4)

```

## Plot PCA
```{r}

cols<-c("#152370","#4bf52a","#f5e107","#f5200c","#fa750f","#18e0f2")

clusters<-cutree(hierarchical.cluster, k=6)
fviz_cluster(list(data = data_clust2, cluster = clusters),
            ellipse.type = "none",stand = FALSE,
             palette = cols, ggtheme = theme_classic()) +
             #scale_shape_manual(values=c("16","16","16","16","16","16"))
            scale_shape_manual(values=c(19,19,19,19,19,19)) +
            ggtitle("") +
            theme(axis.title = element_text(size=20, hjust=0.5), axis.ticks.x=element_blank()) +
            theme(axis.text=element_text(size=16), legend.position="none")

#ggsave("pca.png", path="~/Downloads", width = 4.5, height = 4)

```
## Assess cluster stats 
```{r}

cluster.stats(distance.matrix, data_ap$clusters)

```

## Plot silhouette statistics 
Silhouette value is a measure of how similar an object is to its own cluster (cohesion) compared to other clusters (separation), it ranges from -1 to 1

High numbers indicate object is well matched to its own cluster and poorly matched to neighboring clusters 

Negative means probably belongs in a different cluster 

```{r}

## calculate silhouette
sil <- silhouette(data_ap$clusters, distance.matrix)

# Summary of silhouette analysis
si.sum <- summary(sil)

# Average silhouette width of each cluster
si.sum$clus.avg.widths

neg_sil_index<-which(sil[,'sil_width']<0)
sil[neg_sil_index, ,drop=FALSE]

## plot silhouettes
cols<-c("#152370","#4bf52a","#f5e107","#f5200c","#fa750f","#18e0f2")
p<-fviz_silhouette(sil, print.summary = F) + 
  scale_colour_manual(values = cols) +
  scale_fill_manual(values = cols,  labels=c("cingulo-opercular","dorsal frontal parietal", "mPFC-STS", "parahippocampal", "medial parietal", "entorhinal")) + 
  theme_cowplot() +
  guides(col=FALSE) +
  theme(plot.title = element_text(size=24, face="bold", hjust=0.5), axis.ticks.x=element_blank(), legend.position="none") +
  theme(axis.title = element_text(size=20, hjust=0.5), axis.ticks.x=element_blank()) +
  theme(axis.text.y=element_text(size=16),axis.text.x=element_text(size=12, angle=90) ) +
  labs(x="", y="silhouette width")
p$layers[[2]]$aes_params$colour <- "black"
p

#ggsave("sil.png", path="~/Downloads", width = 6.5, height = 4)

```





