---
title: "Plot individual cluster profiles comprising the superclusters"
output: html_notebook
---

Version 1.0, July 2025, SA

This script plots the connectivity profiles for individual clusters that comprise the superclusters. See "Neocortical Cluster Label" column of Table S1 for information on cluster numbers. 

Input: Hippo_BinxAxis_F_betas.txt

Output: individual cluster profile for whichever desired cluster. Figs S5-10 D panels.

# Packages and functions
```{r}

library(nlme)
library(ggplot2)
library(emmeans)

```

# Read in and format the data 
```{r}
## read in the data 
dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos"
data <- read.table(paste(dir, "/GitHub/Hippo_BinxAxis_F_betas.txt", sep=""), header = TRUE) 

## collapse over hem 
data_ag_hem<-aggregate(data[, "beta"], by=(list(data$subj, data$ax, data$bin, data$clust)), mean)
colnames(data_ag_hem)<-c("subj","ax","bin","clust","beta")

## separate into ant and post and subtract 
data_ag_hem_ant<-subset(data_ag_hem, ax=="ant")
data_ag_hem_post<-subset(data_ag_hem, ax=="post")
ant_post<-data_ag_hem_ant$beta-data_ag_hem_post$beta

## make new dataset 
data_ap<-cbind(data_ag_hem_ant, ant_post)
data_ap<-subset(data_ap, select=-c(ax,beta)) 

```

# Plot 
```{r}

## specify which subcluster you want to look at 
data_clust<-subset(data_ap, clust=="26") ### CHANGE NUMBER OF CLUSTER HERE 

## model 
model<-lme(ant_post ~ as.factor(bin), random= ~1|subj, data=data_clust, na.action=na.omit)

## extract emmean and se 
plot_summary <-emmeans(model, specs = c("bin")) #se is between subjects here 
stats <- as.data.frame(plot_summary)

## plot 
ggplot(stats, aes(x=bin, y=emmean)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower.CL, ymax=upper.CL))+
  labs(x="", y="ant-post connectivity") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.1,0.1), breaks=seq(-0.1,0.1,.02))+
  geom_hline(yintercept=0)+
  ggtitle("29")+
  theme(plot.title = element_text(size=18, face="bold", hjust=0.5))


```