---
title: "Hemisphere x Long-Axis x Age Bin Interactions"
output: html_notebook
---

Version 1.0, July 2025, SA

We ran a linear mixed model predicting hippocampal-voxelwise whole brain connectivity as a function of hemisphere, long-axis, and age bin. We identified neocortical clusters where there was a significant interaction between hemisphere, long-axis, and age bin. This script plots hippocampal connectivity with each of these significant clusters. 

Input: Hippo_BinxHemxAxis_F_betas.txt

Output: Plots in Fig S2C

# Packages and functions
```{r}

library(nlme)
library(ggplot2)
library(emmeans)
library(Rmisc)

```

# Read in the data 
```{r}

dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos"
data <- read.table(paste(dir, "/GitHub/Hippo_BinxHemxAxis_F_betas.txt", sep=""), header = TRUE) 
data <- subset(data, clust != 5) ## when threshold out clusters <10 voxels, top 4 survive 

```

# Plot each cluster
## cluster 1 
```{r}

## specify cluster 
data_clust<-subset(data, clust=="1")

## colours 
clrs<-c('#fa1111','#fa5757','#fc9a9a','#fcc2c2') ### cluster 1 

## identify cluster number for title in plot below 
clust_num<- data_clust[1,"clust"]

## model 
model<-lme(beta ~ ax*hem*bin, random= ~1|subj, data=data_clust, na.action=na.omit)

## prepare to plot 
## use estimated marginal melans 
plot_summary <-emmeans(model, specs = c("ax","hem","bin")) #se is not for within subject data here
plot_summary_df <- as.data.frame(plot_summary)
## use within subject se 
stats<-summarySEwithin(data=data_clust, measurevar = "beta", withinvars = c("bin","hem","ax"), idvar = "subj", na.rm=TRUE)
## add estiamted marginal mean to the stats dataset 
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta, fill=ax:hem)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_fill_manual(values=clrs, labels = c("L ant", "R ant", "L post", "R post"))+ ### mPFC-STS
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(0,0.7), breaks=seq(0,0.7,.2))+
  geom_hline(yintercept=0)+
  ggtitle(clust_num)+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom", legend.text=element_text(size=14))

#ggsave("temp.png", path="~/Downloads", width = 4.5, height = 4.5)


```

## cluster 2 
```{r}

## specify cluster 
data_clust<-subset(data, clust=="2")

## colours 
clrs<-c('#02ebf7','#6ff1f7','#a4f9fc','#d2fbfc') ### cluster 2

## identify cluster number for title in plot below 
clust_num<- data_clust[1,"clust"]

## model 
model<-lme(beta ~ ax*hem*bin, random= ~1|subj, data=data_clust, na.action=na.omit)

## prepare to plot 
## use estimated marginal melans 
plot_summary <-emmeans(model, specs = c("ax","hem","bin")) #se is not for within subject data here
plot_summary_df <- as.data.frame(plot_summary)
## use within subject se 
stats<-summarySEwithin(data=data_clust, measurevar = "beta", withinvars = c("bin","hem","ax"), idvar = "subj", na.rm=TRUE)
## add estiamted marginal mean to the stats dataset 
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta, fill=ax:hem)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_fill_manual(values=clrs, labels = c("L ant", "R ant", "L post", "R post"))+ ### mPFC-STS
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(0,0.7), breaks=seq(0,0.7,.2))+
  geom_hline(yintercept=0)+
  ggtitle(clust_num)+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom", legend.text=element_text(size=14))

#ggsave("temp.png", path="~/Downloads", width = 4.5, height = 4.5)


```

## cluster 3 
```{r}

## specify cluster 
data_clust<-subset(data, clust=="3")

## colours 
clrs<-c('#02fa13','#63ff6e','#96ff9e','#d6ffda') ### cluster 3 
#clrs<-c('#fc00ec','#ff6bf6','#ff94f8','#fcbbf8') ### cluster 4

## identify cluster number for title in plot below 
clust_num<- data_clust[1,"clust"]

## model 
model<-lme(beta ~ ax*hem*bin, random= ~1|subj, data=data_clust, na.action=na.omit)

## prepare to plot 
## use estimated marginal melans 
plot_summary <-emmeans(model, specs = c("ax","hem","bin")) #se is not for within subject data here
plot_summary_df <- as.data.frame(plot_summary)
## use within subject se 
stats<-summarySEwithin(data=data_clust, measurevar = "beta", withinvars = c("bin","hem","ax"), idvar = "subj", na.rm=TRUE)
## add estiamted marginal mean to the stats dataset 
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta, fill=ax:hem)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_fill_manual(values=clrs, labels = c("L ant", "R ant", "L post", "R post"))+ ### mPFC-STS
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(0,0.7), breaks=seq(0,0.7,.2))+
  geom_hline(yintercept=0)+
  ggtitle(clust_num)+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom", legend.text=element_text(size=14))

#ggsave("temp.png", path="~/Downloads", width = 4.5, height = 4.5)


```

## cluster 4 
```{r}

## specify cluster 
data_clust<-subset(data, clust=="4")

## colours 
clrs<-c('#fc00ec','#ff6bf6','#ff94f8','#fcbbf8') ### cluster 4

## identify cluster number for title in plot below 
clust_num<- data_clust[1,"clust"]

## model 
model<-lme(beta ~ ax*hem*bin, random= ~1|subj, data=data_clust, na.action=na.omit)

## prepare to plot 
## use estimated marginal melans 
plot_summary <-emmeans(model, specs = c("ax","hem","bin")) #se is not for within subject data here
plot_summary_df <- as.data.frame(plot_summary)
## use within subject se 
stats<-summarySEwithin(data=data_clust, measurevar = "beta", withinvars = c("bin","hem","ax"), idvar = "subj", na.rm=TRUE)
## add estiamted marginal mean to the stats dataset 
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta, fill=ax:hem)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_fill_manual(values=clrs, labels = c("L ant", "R ant", "L post", "R post"))+ ### mPFC-STS
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(0,0.7), breaks=seq(0,0.7,.2))+
  geom_hline(yintercept=0)+
  ggtitle(clust_num)+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom", legend.text=element_text(size=14))

#ggsave("temp.png", path="~/Downloads", width = 4.5, height = 4.5)


```
