---
title: "Average anterior-posterior difference plots"
output: html_notebook
---

Version 1.0, July 2025, SA

This script takes matrices from ClustRobust.m, which are adjacency matrix that show where clustering is consistent across people. It reads this information in and selectively averages together the anterior-posterior difference in connectivity to each supercluster based on where there is 100% clustering consistency. 

This is different from R_supercluster_stats.Rmd which extracts for anterior and posterior connectivity separately rather than the difference in anterior psoterior connectivity. 

The script then plots the average anterior-posterior difference in connectivity for each supercluster and age bin, which are the plots in Fig. S5C, Fig. S6C, Fig. S7C, Fig. S8B, Fig S9B, and Fig. S10B

Input: threshMat.csv and Hippo_BinxAxis_F_betas.txt 

# Packages and functions
```{r}

library(tidyr)
library(dplyr)
library(nlme)
library(Rmisc)
library(ggplot2)

```

# Selective averaging the ant-post difference within each supercluster 

```{r}

#### read in binarized, thresholded matrix. 1 is where clustering was consistent across all subjects, 0 was inconsistent 
dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos"
data_threshMat <- read.csv(paste(dir, "/GitHub/threshMat.csv", sep=""), header = FALSE) 

### get rid of diagonal and lower triangle 
diag(data_threshMat)<-0
data_threshMat[lower.tri(data_threshMat)]<-0

### read in our original data and format it for our purposes 
dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos"
data <- read.table(paste(dir, "/GitHub/Hippo_BinxAxis_F_betas.txt", sep=""), header = TRUE) 
### use only top 44 for cluster threshold of 10
data<-subset(data, clust < 45)

## collapse over hem for each subj 
data_ag_hem<-aggregate(data[, "beta"], by=(list(data$subj, data$ax, data$bin, data$clust)), mean)
colnames(data_ag_hem)<-c("subj","ax","bin","clust","beta")

## separate into ant and post and subtract 
data_ag_hem_ant<-subset(data_ag_hem, ax=="ant")
data_ag_hem_post<-subset(data_ag_hem, ax=="post")
ant_post<-data_ag_hem_ant$beta-data_ag_hem_post$beta

## make new dataset of ap difference collapsed over hem for each subj 
data_ap2<-cbind(data_ag_hem_ant, ant_post)
data_ap2<-subset(data_ap2, select=-c(ax,beta)) 

#### specify which rois go into each supercluster and append to our ap2 data 
supercluster_1<-c(1,6,7,28,34,36,41,43)
supercluster_2<-c(2,3,9,11,13,14,16,19,20,22,24,30,31,32,37,38,40,44)
supercluster_3<-c(4,15,17,23,25,26,29,33,35,39)
supercluster_4<-c(5,10)
supercluster_5<-c(8,12,27,42)
supercluster_6<-c(18,21)

data_ap2<-mutate(data_ap2, supercluster = case_when(clust %in% supercluster_1 ~ 1,
                                                    clust %in% supercluster_2 ~ 2,
                                                    clust %in% supercluster_3 ~ 3,
                                                    clust %in% supercluster_4 ~ 4,
                                                    clust %in% supercluster_5 ~ 5,
                                                    clust %in% supercluster_6 ~ 6))

### initialize some empty lists to append to later, for each supercluster 
data_sc_1<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_1)<-c("subj","bin","superclust","beta")

data_sc_2<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_2)<-c("subj","bin","superclust","beta")

data_sc_3<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_3)<-c("subj","bin","superclust","beta")

data_sc_4<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_4)<-c("subj","bin","superclust","beta")

data_sc_5<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_5)<-c("subj","bin","superclust","beta")

data_sc_6<-data.frame(matrix(nrow=0, ncol=5))
colnames(data_sc_6)<-c("subj","bin","superclust","beta")


### loop through matrix and when there is a 1, average the two roi hippo-cortical conn values together 
for (r in 1:nrow(data_threshMat)) { 
  
  for (c in 1:ncol(data_threshMat)) {
    
    if (data_threshMat[r,c]==1){
      
      #print("yes")
      
      temp_subset<-subset(data_ap2, (clust==r | clust==c))
      
      ### average across clust 
      temp_agg<-aggregate(temp_subset[, "ant_post"], by=(list(temp_subset$subj,temp_subset$bin, temp_subset$supercluster)), mean)
      colnames(temp_agg)<-c("subj","bin","superclust","beta")
      
      if (temp_agg[1,"superclust"]=="1") {
        data_sc_1<-rbind(data_sc_1, temp_agg)
      }
      
      if (temp_agg[1,"superclust"]=="2") {
        data_sc_2<-rbind(data_sc_2, temp_agg)
      }
      
      if (temp_agg[1,"superclust"]=="3") {
        data_sc_3<-rbind(data_sc_3, temp_agg)
      }
      
      if (temp_agg[1,"superclust"]=="4") {
        data_sc_4<-rbind(data_sc_4, temp_agg)
      }
      
      if (temp_agg[1,"superclust"]=="5") {
        data_sc_5<-rbind(data_sc_5, temp_agg)
      }
      
      if (temp_agg[1,"superclust"]=="6") {
        data_sc_6<-rbind(data_sc_6, temp_agg)
      }
      
    }
    
  }
  
}
  

### average per participant for each supercluster 
data_sc_1<-aggregate(data_sc_1[, "beta"], by=(list(data_sc_1$subj, data_sc_1$bin)), mean)
colnames(data_sc_1)<-c("subj","bin","beta")
data_sc_1$supercluster<-1

data_sc_2<-aggregate(data_sc_2[, "beta"], by=(list(data_sc_2$subj, data_sc_2$bin)), mean)
colnames(data_sc_2)<-c("subj","bin","beta")
data_sc_2$supercluster<-2

data_sc_3<-aggregate(data_sc_3[, "beta"], by=(list(data_sc_3$subj, data_sc_3$bin)), mean)
colnames(data_sc_3)<-c("subj","bin","beta")
data_sc_3$supercluster<-3

## nothing in this one, parahippocampal 
# data_sc_4<-aggregate(data_sc_4[, "beta"], by=(list(data_sc_4$subj, data_sc_4$bin)), mean)
# colnames(data_sc_4)<-c("subj","bin","beta")
# data_sc_4$supercluster<-4

data_sc_5<-aggregate(data_sc_5[, "beta"], by=(list(data_sc_5$subj, data_sc_5$bin)), mean)
colnames(data_sc_5)<-c("subj","bin","beta")
data_sc_5$supercluster<-5

data_sc_6<-aggregate(data_sc_6[, "beta"], by=(list(data_sc_6$subj, data_sc_6$bin)), mean)
colnames(data_sc_6)<-c("subj","bin","beta")
data_sc_6$supercluster<-6


```


# Plotting the average ant-post difference per supercluster

## Cingulo-opercular supercluster
```{r, fig.height=4,fig.width=4.5}

data_SC<-data_sc_1 ### this is weighted averaged data across component clusters 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "beta", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.1,0.1), breaks=seq(-0.1,0.1,.02))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("cingulo-operc.png", path="~/Downloads", width = 4.5, height = 4.5)

```

## Dorsal frontal parietal supercluster
```{r, fig.height=4,fig.width=4.5}


data_SC<-data_sc_2 ### this is weighted averaged data across component clusters 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "beta", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.1,0.1), breaks=seq(-0.1,0.1,.02))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("dorsalFPN.png", path="~/Downloads", width = 4.5, height = 4.5)

```

## mPFC-STS supercluster 
```{r, fig.height=4,fig.width=4.5}

data_SC<-data_sc_3 ### this is weighted averaged data across component clusters 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "beta", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.1,0.1), breaks=seq(-0.1,0.1,.02))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("vmpfc_sts.png", path="~/Downloads", width = 4.5, height = 4.5)

```

## Medial parietal supercluster
```{r, fig.height=4,fig.width=4.5}

data_SC<-data_sc_5 ### this is weighted averaged data across component clusters 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "beta", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.1,0.1), breaks=seq(-0.1,0.1,.02))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("medialparietal.png", path="~/Downloads", width = 4.5, height = 4.5)

```

## Entorhinal supercluster
```{r, fig.height=4,fig.width=4.5}

data_SC<-data_sc_6 ### this is weighted averaged data across component clusters 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "beta", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$beta + stats$se
stats$lower_se<-stats$beta - stats$se

## plot 
ggplot(stats, aes(x=bin, y=beta)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.2,0.2), breaks=seq(-0.2,0.2,.05))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("entorhinal.png", path="~/Downloads", width = 4.5, height = 4.5)

```

## Parahippocampal supercluster
```{r, fig.height=4,fig.width=4.5}

data_SC<-subset(data_ap2, supercluster==4) ### not plotting weighted average here because the two clusters comprising this supercluster were not clustered together 100% of the time. This is the un-weighted data. 

## prepare to plot 
stats<-summarySE(data=data_SC, measurevar = "ant_post", groupvars = c("bin"), na.rm=TRUE)
stats$upper_se<-stats$ant_post + stats$se
stats$lower_se<-stats$ant_post - stats$se

## plot 
ggplot(stats, aes(x=bin, y=ant_post)) + 
  theme_classic() +
  geom_bar(stat = "identity", position = "dodge", width = 0.8, fill="steelblue") + 
  geom_errorbar(width = .0, position = position_dodge(0.8), aes(ymin=lower_se, ymax=upper_se))+
  labs(x="", y="ant-connectivity (Fisher z)") +
  theme(axis.title = element_text(size=20), axis.ticks.x=element_blank()) +
  theme(axis.text=element_text(size=14))+
  labs(fill="") +
  scale_x_discrete(limits=c("1", "2", "3", "4"), labels=c("0-6mo", "7-12mo", "13-18mo","19-25mo"))+
  scale_y_continuous(expand = c(0, 0), limits = c(-0.2,0.2), breaks=seq(-0.2,0.2,.05))+
  geom_hline(yintercept=0)+
  ggtitle(" ")+
  theme(plot.title = element_text(size=22, face="bold", hjust=0.5), legend.position="bottom")

#ggsave("parahippocampal.png", path="~/Downloads", width = 4.5, height = 4.5)

```




