---
title: "tSNR analyses"
output: html_notebook
---

Version 1.0, July 2025, SA

This script plots mean and standard deviation of connectivity across the age bins examined. 

Input: Mean_and_SD_cortical_conn_Schaefer2018_200Parcels.csv

Output: Figs S12 and S13

# Read in packages 
```{r}

library(ggplot2)
library(dplyr)

```

# Read in and prepare the data 
```{r}
## Read in the data 
dir<-"/Users/audrainsp/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/BabyHippos/R_BabyHippos/GitHub"
data<- read.csv(paste(dir, "/Mean_and_SD_cortical_conn_Schaefer2018_200Parcels.csv", sep=""), header = TRUE) 

## subset appropriate age ranges 
data<-subset(data, (age<26 | age>35) )

attach(data)

```

# Without regressing out motion 

## Variability of connectivity 
```{r, fig.height=4,fig.width=7}

## correlation
cor.test(age,sd)

## Plot 
ggplot(data, aes(y=sd, x=age)) + 
  geom_vline(xintercept = 6, color="grey", linetype="solid") +
  geom_vline(xintercept = 12, color="grey", linetype="solid") +
  geom_vline(xintercept = 18, color="grey", linetype="solid") +
  geom_vline(xintercept = 25, color="grey", linetype="solid") +
  geom_point(size=2, alpha=0.7) + 
  geom_smooth(method="loess", level=0.95, col="black") + 
  theme_classic() +
  theme(text = element_text(size = 18), axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), plot.title = element_text(face="bold", hjust=0.5, size=22), legend.position="none")+
  scale_shape_manual(values=c(20))+
  guides(shape = FALSE, size = FALSE)+
  #scale_y_continuous(expand = c(0, 0), limits = c(0,0.3), breaks=seq(0,0.3,.05))+
  scale_x_continuous(expand = c(0, 0), limits = c(0,70), breaks=seq(0,70,5))+
  scale_y_continuous(limits = c(0.28,0.38), breaks=seq(0.28,0.38,.01))+
  labs(y = "standard deviation
       ",
       x = "age (months)",
       title = "")

#ggsave("sd_5bins.png", path="~/Downloads", width = 7, height = 4)

```
## Mean connectivity 
```{r, fig.height=4,fig.width=7}

## correlation
cor.test(age,mean)

## plot 
ggplot(data, aes(y=mean, x=age)) + 
  geom_vline(xintercept = 6, color="grey", linetype="solid") +
  geom_vline(xintercept = 12, color="grey", linetype="solid") +
  geom_vline(xintercept = 18, color="grey", linetype="solid") +
  geom_vline(xintercept = 25, color="grey", linetype="solid") +
  geom_point(size=2, alpha=0.7) + 
  geom_smooth(method="loess", level=0.95, col="black") + 
  theme_classic() +
  theme(text = element_text(size = 18), axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), plot.title = element_text(face="bold", hjust=0.5, size=22), legend.position="none")+
  scale_shape_manual(values=c(20))+
  guides(shape = FALSE, size = FALSE)+
  scale_x_continuous(expand = c(0, 0), limits = c(0,70), breaks=seq(0,70,5))+
  scale_y_continuous(limits = c(0,0.6), breaks=seq(0,0.6,.1))+
  labs(y = "mean connectivity
       ",
       x = "age (months)",
       title = "")

#ggsave("mean_5bins.png", path="~/Downloads", width = 7, height = 4)


```

# After regressing out motion

## Variability of connectivity 
```{r, fig.height=4,fig.width=7}

m1<-lm(sd~mean_DiffMag, data=data)
mean_sd<-mean(data$sd) ## to add back in so scale makes sense 
#resid(m1)

ggplot(data, aes(y=resid(m1)+mean_sd, x=age)) + 
  geom_vline(xintercept = 6, color="grey", linetype="solid") +
  geom_vline(xintercept = 12, color="grey", linetype="solid") +
  geom_vline(xintercept = 18, color="grey", linetype="solid") +
  geom_vline(xintercept = 25, color="grey", linetype="solid") +
  geom_point(size=2, alpha=0.7) + 
  geom_smooth(method="loess", level=0.95, col="black") + 
  theme_classic() +
  theme(text = element_text(size = 18), axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), plot.title = element_text(face="bold", hjust=0.5, size=22), legend.position="none")+
  scale_shape_manual(values=c(20))+
  guides(shape = FALSE, size = FALSE)+
  scale_x_continuous(expand = c(0, 0), limits = c(0,70), breaks=seq(0,70,5))+
  scale_y_continuous(limits = c(0.28,0.38), breaks=seq(0.28,0.38,.01))+
  labs(y = "standard deviation
       (controlling for mean motion)",
       x = "age (months)",
       title = "")

#ggsave("sd_5bins_nomotion.png", path="~/Downloads", width = 7, height = 4)
```

## Mean connectivity 
```{r, fig.height=4,fig.width=7}
m1<-lm(mean~mean_DiffMag, data=data)
mean_mean<-mean(data$mean) ## to add back in so scale makes sense 

ggplot(data, aes(y=resid(m1)+mean_mean, x=age)) + 
  geom_vline(xintercept = 6, color="grey", linetype="solid") +
  geom_vline(xintercept = 12, color="grey", linetype="solid") +
  geom_vline(xintercept = 18, color="grey", linetype="solid") +
  geom_vline(xintercept = 25, color="grey", linetype="solid") +
  geom_point(size=2, alpha=0.7) +  
  geom_smooth(method="loess", level=0.95, col="black") + 
  theme_classic() +
  theme(text = element_text(size = 18), axis.text.x = element_text(size=18), axis.text.y = element_text(size=18), plot.title = element_text(face="bold", hjust=0.5, size=22), legend.position="none")+
  scale_shape_manual(values=c(20))+
  guides(shape = FALSE, size = FALSE)+
  scale_x_continuous(expand = c(0, 0), limits = c(0,70), breaks=seq(0,70,5))+
  scale_y_continuous(limits = c(0,0.6), breaks=seq(0,0.6,.1))+
  labs(y = "mean connectivity 
       (controlling for mean motion)",
       x = "age (months)",
       title = "")

#ggsave("mean_5bins_nomotion.png", path="~/Downloads", width = 7, height = 4)

```

# Assessing mean conn inside and outside of infantile amnesia window 

```{r}

## t test of mean connectivity in participants within and outside the infantile amnesia window 
## create lable for AI and A
data<-data %>%
  mutate(amnesia=case_when(age<26 ~ "inside",
                      age>35 ~ "outside"))
data_restricted<-subset(data, (age>18 & age<26) | age>35)

## mean conn 
m1<-lm(mean~mean_DiffMag, data=data_restricted)
t.test(resid(m1) ~ data_restricted$amnesia)


```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

